---
title: "Exploring the Iris Dataset with Plotly"
author: "K Hemant"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

## Introduction

Welcome to the interactive exploration of the classic **Iris dataset**!  
In this presentation, we'll use **Plotly** for dynamic visualizations.

---

## About the Iris Dataset

- Contains measurements for **3 species** of iris flowers.
- Variables:
  - Sepal Length & Width
  - Petal Length & Width
- Famous for classification tasks and visual analytics.

---

## Required Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(plotly)


## Preview of the Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

head(iris)


## Scatter Plot: Sepal.Length vs Petal.Length

plot_ly(iris, x = ~Sepal.Length, y = ~Petal.Length,
        color = ~Species, type = 'scatter', mode = 'markers',
        marker = list(size = 10, opacity = 0.7),
        text = ~paste("Sepal Width:", Sepal.Width, "<br>Petal Width:", Petal.Width))

## Histogram: Sepal Length Distribution
plot_ly(iris, x = ~Sepal.Length, type = 'histogram',
        nbinsx = 20, color = ~Species, barmode = "overlay",
        opacity = 0.6)

## Boxplot: Petal Width by Species

plot_ly(iris, y = ~Petal.Width, x = ~Species, type = "box", color = ~Species)

## Scatter Plot: Sepal.Width vs Petal.Width
plot_ly(iris, x = ~Sepal.Width, y = ~Petal.Width,
        color = ~Species, type = 'scatter', mode = 'markers',
        marker = list(size = 10, opacity = 0.7))


## Pairwise Feature Scatter Matrix
plot_ly(data = iris, type = 'splom',
        dimensions = list(
            list(label = 'Sepal Length', values = ~Sepal.Length),
            list(label = 'Sepal Width', values = ~Sepal.Width),
            list(label = 'Petal Length', values = ~Petal.Length),
            list(label = 'Petal Width', values = ~Petal.Width)
        ),
        text = ~Species,
        marker = list(color = ~as.integer(Species),
                      colorscale = 'Viridis',
                      showscale = FALSE,
                      line = list(width = 0.5, color = 'black')))

## Correlation Heatmap
numeric_iris <- iris %>% select(-Species)
corr <- round(cor(numeric_iris), 2)

plot_ly(
  x = colnames(corr),
  y = rownames(corr),
  z = corr,
  type = "heatmap",
  colorscale = "Blues"
)


## Key Takeaways

* Setosa is distinctly different in petal features
* Petal.Length and Petal.Width are strongly correlated
* Plotly makes EDA fun and interactive

---






