---
title: "Exploring the Iris Dataset with Plotly"
author: "K Hemant"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

## Introduction

Welcome to the interactive exploration of the classic **Iris dataset**!\
In this presentation, we'll use **Plotly** for dynamic visualizations.

------------------------------------------------------------------------

## About the Iris Dataset

-   Measurements of 150 iris flowers\
-   3 species: *setosa*, *versicolor*, *virginica*\
-   Features:
    -   Sepal Length & Width\
    -   Petal Length & Width

------------------------------------------------------------------------

## Required Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(plotly)
```

------------------------------------------------------------------------

## Preview of the Data

```{r}
head(iris)
```

------------------------------------------------------------------------

## Scatter Plot: Sepal.Length vs Petal.Length

```{r}
plot_ly(iris, x = ~Sepal.Length, y = ~Petal.Length,
        color = ~Species, type = 'scatter', mode = 'markers',
        marker = list(size = 10, opacity = 0.7),
        text = ~paste("Sepal Width:", Sepal.Width, "<br>Petal Width:", Petal.Width))
```

------------------------------------------------------------------------

## Histogram: Sepal Length Distribution

```{r}
plot_ly(iris, x = ~Sepal.Length, type = 'histogram',
        nbinsx = 20, color = ~Species, barmode = "overlay", opacity = 0.6)
```

------------------------------------------------------------------------

## Boxplot: Petal Width by Species

```{r}
plot_ly(iris, y = ~Petal.Width, x = ~Species, type = "box", color = ~Species)
```

------------------------------------------------------------------------

## Scatter Plot: Sepal.Width vs Petal.Width

```{r}
plot_ly(iris, x = ~Sepal.Width, y = ~Petal.Width,
        color = ~Species, type = 'scatter', mode = 'markers',
        marker = list(size = 10, opacity = 0.7))
```

------------------------------------------------------------------------

## Pairwise Feature Scatter Matrix

```{r}
plot_ly(data = iris, type = 'splom',
        dimensions = list(
            list(label = 'Sepal Length', values = ~Sepal.Length),
            list(label = 'Sepal Width', values = ~Sepal.Width),
            list(label = 'Petal Length', values = ~Petal.Length),
            list(label = 'Petal Width', values = ~Petal.Width)
        ),
        text = ~Species,
        marker = list(color = ~as.integer(Species),
                      colorscale = 'Viridis',
                      showscale = FALSE,
                      line = list(width = 0.5, color = 'black')))
```

------------------------------------------------------------------------

## Correlation Heatmap

```{r}
numeric_iris <- iris %>% select(-Species)
corr <- round(cor(numeric_iris), 2)

plot_ly(
  x = colnames(corr),
  y = rownames(corr),
  z = corr,
  type = "heatmap",
  colorscale = "Blues"
)
```

------------------------------------------------------------------------

## Key Takeaways

-   Setosa is distinctly different in petal features\
-   Petal.Length and Petal.Width are strongly correlated\
-   Plotly makes EDA fun and interactive\
-   Interactive plots = better insight & presentation

------------------------------------------------------------------------
